\documentclass[a4paper,titlepage]{article}
\usepackage{url}
\title{Nigori: Storing Secrets in the Cloud}
\author{Ben Laurie\\
(benl@google.com)}

\begin{document}
\maketitle

\def\qe#1{\begin{equation}#1\end{equation}}

\section{Introduction}

Secure login is something we would clearly like, but achieving it
practically for the majority users turns out to be surprisingly
hard. In particular, we would like it to be possible for a user to log
in from anywhere, using nothing except standard software and
information that is in his head, without making himself vulnerable to
phishing attacks. We, of course, do not exclude more stringent
requirements for login, but we wish to satisfy this lowest common
denominator.

We start by listing detailed requirements (and non-requirements) for a
practical system and then present a system which satisfies those
requirements,

\section{Requirements}

\begin{enumerate}

  \item The user should not have to remember any more than a single
    ``master'' password and a URL.

  \item The user should be able to log in to any site from any machine
    with appropriate software installed.

  \item We should not require sites at which the user logs in (relying
    parties or RPs) to change how users log in. That is, the user
    should still be able to be logged in with a password.

  \item For sites that want to support stronger authentication
    mechanisms, such as public/private keypairs or other cryptographic
    mechanisms, this option should be available.

  \item When a user logs in at an RP with a password, it should be
    different from the password used at any other RP.

  \item It should be possible to change the password used at any
    particular RP.

  \item The user should not have to trust any single external entity
    in order to remain secure.

  \item The user should have choice over which software implements the
    system (i.e. we believe in open standards).

  \item Attacks on master passwords must not be offline attacks,
    except for attacks by any entity (or group of entities) which are
    providing the storage service. Clearly, they cannot be prevented
    from mounting an ``offline'' attack by simulating the user in a
    closed network environment.

  \item Modification of stored secrets should be detectable.

  \item The user should be able to use more than one device.

\end{enumerate}

\section{Non-requirements}

\begin{enumerate}

  \item A standard browser or other client as it exists today - we
    explicitly do not exclude the possibility of a strong solution
    requiring modification to existing browsers and
    clients. Nevertheless, we will present a variant of the system
    which should work in existing browsers.

\end{enumerate}

\section{Consequences of the Requirements}

It seems clear that in order to satisfy the requirements there must be
a place or places where a user's secrets (passwords, private keys,
etc) are stored. These secret stores must not, however, be able to
access those secrets.

This means the secrets need to be encrypted in such a way that the
store has no access to the encryption key.

Because attacks on the master password should not be offline, the user
must authenticate to the store before retrieving the secrets, and that
authentication must be derived from the master password.

This, in turn, means that the user cannot authenticate by simply
providing his password to the store, as is currently standard with web
authentication.

The secret stores must also be accessible from anywhere - that is,
they need to be ``in the cloud''.

Because there should be no offline attacks available, and
modifications should be detectable, only the legitimate user should
ever be able to see the secrets. This means that the secrets should be
retrieved over a secure channel, otherwise a man-in-the-middle or a
snooper would be able to take the encrypted secret and try master
passwords until the integrity check succeeded. We do not discuss in
any detail what secure channel is used because these already exist,
for example in the form of TLS or OpenSSH connections between client
and server.

\section{Details of the System}

Here we describe the Nigori protocols in some detail (precise
encodings and parameters, however, are described in the Nigori
protocol document).

All connections are assumed to be over a secure channel which can
prevent eavesdropping and man-in-the-middle attacks, for example a TLS
connection authenticated by a server certificate.

\subsection{Key Derivation}

We allow two mechanisms for key derivation - the first reduces the
level of trust the user must have in the service providers, but
provides no mechanisms for recovery if the user forgets his password
(other than, obviously, the user providing a way to recover the
password itself, such as, for example, writing it down, or telling it
to a trusted friend). The second allows for recovery in the event that
the user forgets the password, in exchange for greater exposure to
service providers. We call these unassisted and assisted
password-based key derivation, respectively.

In both cases, four keys are derived, $K_{user}$, used to
authenticate the user to service providers, $K_{enc}$, used to encrypt
secrets, $K_{mac}$, used to check the integrity of encrypted
secrets (that is, to allow the detection of accidental or malicious
tampering) and $K_{iv}$ used to generate intialisation vectors and
prevent common prefix attacks.

\subsubsection{Unassisted Password-based Key Derivation}


In the unassisted mode, $K_{user}$, $K_{enc}$, $K_{mac}$ and $K_{iv}$ are all derived from the master password in such a way that they are not derivable from each other.

For example, in the current protocol, we use PBKDF2 [Kaliski, B
  (2000), ``RFC 2898: PKCS \#5: Password-Based Cryptography
  Specification Version 2.0'', http://tools.ietf.org/html/rfc2898]
with a different count for each secret.

\subsubsection{Assisted Password-based Key Derivation}

In the assisted mode, first we retrieve a key $K_{master}$ for the
user from some server other than the storage server (we could call
this the recovery server). It is up to that server to decide how to
create this key, but it must be able to recreate the same key even
after the user has forgotten his password for the recovery server (or
whatever other means he uses to log in to it). For example, the
recovery server could derive $K_{master}$ from the hash of the user's
password that it presumably stores to use in password verification.

The client then derives $K_{enc}$, $K_{mac}$ amd $K_{iv}$ from $K_{master}$, and
$K_{user}$ from the user's master password (which may be the same as
the password he uses to log in to the recovery server). Note that even
if both passwords are the same, an honest recovery server cannot
derive $K_{user}$ from data it has stored, because it should not store
the password, just something derived from it in order to verify the
user.

Recovery from a forgotten password in this case is possible - the user
performs account recovery in the usual way (e.g. through a helpdesk,
via email or SMS, by proving knowledge of various other secrets, etc.)
at the storage and/or recovery servers. The recovery server then
provides the old $K_{master}$ (we recommend that it should change
after account recovery) and this the client can still decrypt the
user's secrets - and re-encrypt with the new $K_{enc}$, $K_{mac}$ and $K_{iv}$.

\subsection{Storage}

Now that we've dealt with the derivation of the four keys, storage
and retrieval of secrets is relatively straightforward.

Firstly, the user authenticates to the storage server using
$K_{user}$, then either sends or retrieves secrets which are encrypted
and authenticated with $K_{enc}$, $K_{mac}$ and $K_{iv}$.

There are a wide variety of ways the user could authenticate using
$K_{user}$, but we have chosen to use a DSA signature (with $K_{user}$
as the private key) over the username, the server name, a timestamp
and a nonce. The server checks that all of these are valid, that the
timestamp is recent, that the particular timestamp/nonce combination
has never been seen before and that the signature is valid. Note that
the timestamp is needed to limit the size of the database used to
prevent replays.

The advantage of a DSA signature for this purpose is that it does not
give an attacker who has access to the communications (for example,
one who has taken over the storage server, or a storage server that is
evil) any kind of password equivalent.

Regardless of authentication mechanism used, because $K_{enc}$ cannot
be derived from $K_{user}$ (other than through a brute force attack on
the user's password), the storage server is unable to decrypt the
user's secrets.

\subsection{Split Storage}

Once all of the above is in place, the user could choose to have his
client store his secrets split across multiple servers, such that to
even mount an attack on the secrets multiple storage servers would
have to collude (note, however, that this is likely to be a poor
defence unless he uses different passwords at different servers).

In addition to splitting for security reasons, the user might choose
to use multiple servers for resilience.

When splitting for security, a threshold encryption scheme could be
used, such as Shamir's secret splitting scheme [Shamir, Adi (1979),
  ``How to share a secret'', Communications of the ACM 22 (11):
  612613]. When splitting for resilience obviously each server could
store the whole of the secret, but nevertheless the user might choose
to use a split to mitigate the risks involved (for example, a user who
splits for resilience is quite likely to do so because he runs his own
server, but is not confident of 100\% uptime and so wants to use the
servers of friends as backups - but does not want to trust or burden
his friends with the whole of his secrets).

\subsection{Details of Storage}

If we are to achieve our requirements, we must specify not only how we
protect our secrets, but exactly what is stored, in order to allow
interoperable implementations. Therefore the detailed protocol
specifies bit-level formats for the various types of secrets (at this
time, these are passwords [i.e. random blobs], RSA private keys and
the locations of split storage servers).
%TODO(drt24) this is not how we have actually implemented it - we have no concept of different types on the server.

Each secret is stored against a index (I hesitate to use the
traditional word ``key'' at this point!) which is the encryption of
the information needed to use the resource (e.g. the URL at which the
password is entered) and its type.

We also store all the old values of encrypted secrets in case of
failed password change attempts, disaster recovery at the RPs and the
like which might require a fallback to an older version of the
secret. Obviously it would make sense to expire old values eventually.
This is achieved by storing a set of revisions for each index with
a value associated with each revision.

Finally the storage servers should be prepared to list the (encrypted)
indices so that users can manually deal with issues like sites
which change their name or ownership, or sites where login is
available at multiple URLs.

\section{Variations on the Theme}

\subsection{Unmodified Client}

One variation we have considered is the situation where the user has
only an existing browser, with no client-side support for the
protocol. In this case some minor modifications of key derivation
details, combined with some standard Javascript and an OpenID or
OAuth-like flow of redirects and UI could be used to retrieve and
exercise the secrets.

In particular, we would operate in the assisted mode and $K_{master}$
would include a component that was derived from the name of the
RP. When the user arrived at the RP it would initiate a flow like
OpenID or OAuth\footnote{Both protocols work by a series of HTTP
  redirects. Each one takes the user to the next actor in the
  protocol, who may then interact with the user, or may directly take
  action purely based on the redirected URL, and then in its turn
  redirects the user's browser, eventually ending up back at the RP
  with retrieved information in the URL of the final redirect.},
redirecting the user to their IdP. The IdP would supply $K_{master}$,
which would be defended against manipulation by the RP because a
dishonest RP supplying the wrong RP name would then not end up being
redirected to at the end of the redirection chain. Instead of
returning directly to the RP, the IdP would instead bounce via the
storage server, which would add the appropriate encrypted secret to
the redirection URL. Finally, the user's browser would end up back at
the RP with $K_{master}$ and the encrypted secret, which a standard
piece of Javascript\footnote{By ``standard'' we mean that it would be
  largely identical at all RPs and could therefore be easily included
  as a static modification to the login page.} served by the RP would
decrypt and enter into a password box.

The great advantage of this scheme over the standard OpenID flow is
that the RP would not have to make any modification to their login
flow other than including the standard Javascript for doing the
redirection and decryption. The end result would be a password typed
in as if by the user.

Note that a dishonest RP would only be able to retrieve the user's
password for itself, which is not an interesting attack.

Also note that, because of this, if the URL becomes too long because
of the size of encrypted secrets and so forth, the IdP and storage
server could instead hand short codes to the RP which it could then
use to retrieve the full versions to be processed by the Javascript.

\subsection{Granting access via OAuth}

If instead of $K_{enc}$, which is the same across all secrets, we used
a value, $K_v$, derived from both the user's password and the lookup
value, then protocols such as OAuth could be used to grant access to
individual secrets. There are two variants of this.

\subsubsection{Variant 1}

The service provider (SP) is granted access to the encrypted key and
is also given the value of $K_v$ so it can decrypt it (and so the
storage server cannot).

\subsubsection{Variant 2}

The SP is granted access and the storage server decrypts the secret for it. There seems to be little value in this variant, but we mention it for completeness.

\subsection{Exercising the Secret on Behalf of the User}

In this variant, the user may be offline and so unable to use the
secret via his client, in which case an OAuth grant could be made as
above, but rather than the SP getting access to the secret itself, yet
another service uses the key on behalf of the user. The advantage of
this variant is that the user could grant SPs it doesn't much trust
access to, for example, signing keys and could later revoke that
access without necessarily revoking the keys.

\section{Conclusion}

We believe we have presented a protocol which achieves all the goals
stated at the outset of this document, and which is also capable of
extension to solve a number of previously unsolved related problems.

Once such a mechanism is in place it becomes feasible for users to
finally heed the impractical advice they have long been given: use a
different password at each site to avoid phishing. Furthermore, it
allows those passwords to be entirely random, and therefore strong.

Furthermore, the system can also be used to gradually move away from
password-based authentication to stronger mechanisms, such as client
certificates or selective disclosure [Laurie, Ben (2007), ``Selective
  Disclosure'', \url{http://www.links.org/files/selective-disclosure.pdf}].

Finally, the system can be incorporated into most sites with almost no
modification, if so desired, whilst allowing a smooth transition to
more advanced mechanisms.

\end{document}
